---
title: "Plot QC results for each genome/transcriptome used in Orthofinder Run2"
author: "Timothy Stephens"
date: "06/08/2023"
output: 
  html_document:
    keep_md: yes
---

# Setup

Setup R env. Load packages and set default image export formats, size and resolution.
```{r setup}
knitr::opts_chunk$set(echo = TRUE,
                      fig.height = 8, 
                      fig.width = 12, 
                      dev = c("png", "pdf"),
                      dpi = 1000)
library(ggplot2)
library(ggtree)
library(aplot)
library(treeio)
library(tibble)
library(dplyr)
library(reshape2)
library(scales)
library(ggrepel)
library(stringr)
library(glue)
library(ggpubr)
options(scipen = 999) #Prevent scientific notation
```





# Functions

Function to plot BUSCO categories - uses the standard BUSCO color scheme and also plots the percent for each category.
```{r func_plot_BUSCO_percentages}
format_values_1 = function(values){
  out <- c()
  for(value in values) {
    if(value != 0){
      out <- c(out, sprintf("  %1.1f%%", value))
    } else {
      out <- c(out, "")
    }
  }
  return(out)
}

plot_BUSCO_percentages = function(data, plot_title, 
                                  plot.axis.text.y=FALSE,
                                  plot.legend=FALSE,
                                  plot_size_ratio=1, plot_family="sans",
                                  my_colors=c("#56B4E9", "#3492C7", "#F0E442", "#F04442") # Color pallet for BUSCO gene categories.
                                  )
{
  p <- data %>%
    melt(id.var="SampleID") %>%
    rename(BUSCO_categories = variable, Count = value) %>%
    mutate(BUSCO_categories = factor(BUSCO_categories, c("Complete", "Single-copy", "Duplicated", "Fragmented", "Missing"))) %>%
    filter(BUSCO_categories!="Complete") %>%
    filter(BUSCO_categories!="Total") %>%
    ggplot(aes(y = Count, x = SampleID, fill = BUSCO_categories)) +
      geom_bar(position = position_stack(reverse = TRUE),
               stat="identity") +
      coord_flip() +
      # geom_text_repel: https://ggrepel.slowkow.com/articles/examples.html
      geom_text_repel(aes(label = format_values_1(Count)), # Add two spaces in from of text so it centers off the number not number%
                size = rel(5)*plot_size_ratio,
                fontface = "bold",
                max.overlaps = Inf, #Set max.overlaps = Inf to override this behavior and always show all labels, regardless of too many overlaps.
                position = position_stack(vjust=0.5, reverse = TRUE),
                xlim = c(-Inf, Inf), # Set xlim or ylim to Inf or -Inf to disable repulsion away from the edges of the panel.
                ylim = c(-Inf, Inf),
                point.size = NA, #size of each point for each text label; Set point.size = NA to prevent label repulsion away from data points.
                force=0.5, # force of repulsion between overlapping text labels
                box.padding=0, # padding around the text label
                direction="y") + # move text labels “both” (default), “x”, or “y” directions
      theme_gray(base_size = 8) + 
      scale_y_continuous(labels = c("0","20","40","60","80","100"), 
                         breaks = c(0,20,40,60,80,100)) +
      scale_fill_manual(values = my_colors,
                        labels = c("Complete (C) and single-copy (S)",
                                   "Complete (C) and duplicated (D)",
                                   "Fragmented (F)",
                                   "Missing (M)")) +
      ggtitle(plot_title) +
      xlab("") +
      ylab("%BUSCOs") +
      theme(plot.title = element_text(family = plot_family,
                                      hjust=0.5, 
                                      colour = "black", 
                                      size = rel(2.2)*plot_size_ratio,
                                      face = "bold")) +
      theme(plot.margin=unit(c(0, 0, 0, 0), "mm")) +
      theme(panel.background = element_rect(color = "#FFFFFF", fill = "white")) +
      theme(panel.grid.minor = element_blank()) +
      theme(panel.grid.major = element_blank()) +
      theme(axis.text.y = element_text(family = plot_family, face = "bold", size = rel(1.66)*plot_size_ratio, colour = "black")) +
      theme(axis.text.x = element_text(family = plot_family, face = "bold", size = rel(1.66)*plot_size_ratio, colour = "black")) +
      theme(axis.line = element_line(linewidth = 1*plot_size_ratio, colour = "black")) +
      theme(axis.ticks.length = unit(0.85*plot_size_ratio, "cm")) +
      theme(axis.ticks.y = element_line(colour = "white", linewidth = 0)) +
      theme(axis.ticks.x = element_line(colour = "#222222")) +
      theme(axis.ticks.length = unit(0.4*plot_size_ratio, "cm")) + 
      theme(axis.title.x = element_text(family = plot_family, size = rel(2)*plot_size_ratio)) +
      theme(legend.position = "bottom", legend.title = element_blank()) + 
      theme(legend.text = element_text(family = plot_family, size = rel(1.2)*plot_size_ratio)) +
      theme(legend.key.size = unit(1.5*plot_size_ratio,"line")) +
      guides(fill = guide_legend(override.aes = list(colour = NULL))) +
      guides(fill = guide_legend(nrow = 1, byrow = TRUE))
  if (!plot.axis.text.y) {
    p <- p + theme(axis.text.y=element_blank())
  }
  if (!plot.legend) {
    p <- p + theme(legend.position = "none")
  }
  return(p)
}
```

Function to plot basic single-category bar chart (not stacked). User provided color for bars.
```{r func_plot_barchart}
format_values_2 = function(values){
  out <- c()
  for(value in values) {
    if(value != 0){
      out <- c(out, scales::comma(value))
    } else {
      out <- c(out, "")
    }
  }
  return(out)
}

plot_barchart = function(data, column.id, 
                         plot_title, xlab, bar.fill, 
                         divide.by=1, y.log2=FALSE, y.log10=FALSE, 
                         add.geom_text=TRUE, plot.axis.text.y=FALSE,
                         plot_size_ratio=1, plot_family="sans")
{
  p <- data %>%
    melt(id.var="SampleID") %>%
    filter(variable==column.id) %>%
    mutate(value = as.numeric(value)) %>%
    mutate(value = value/divide.by) %>%
    ggplot(aes(y = value, x = SampleID)) +
      geom_bar(position = position_stack(),
               stat="identity",
               fill=bar.fill) +
      scale_y_continuous(labels = comma) +
      coord_flip() +
      theme_gray(base_size = 8) + 
      ggtitle(plot_title) +
      xlab("") +
      ylab(xlab) +
      theme(plot.title = element_text(family = plot_family,
                                      hjust=0.5, 
                                      colour = "black", 
                                      size = rel(2.2)*plot_size_ratio,
                                      face = "bold")) +
      theme(plot.margin=unit(c(0, 0, 0, 0), "mm")) +
      theme(legend.position="top",legend.title = element_blank()) + 
      theme(legend.text = element_text(family = plot_family, size = rel(1.2)*plot_size_ratio)) +
      theme(legend.key.size = unit(1.5*plot_size_ratio,"line")) +
      theme(panel.background = element_rect(color = "#FFFFFF", fill = "white")) +
      theme(panel.grid.minor = element_blank()) +
      theme(panel.grid.major = element_blank()) +
      theme(axis.text.y = element_text(family = plot_family, face = "bold", size = rel(1.66)*plot_size_ratio, colour = "black")) +
      theme(axis.text.x = element_text(family = plot_family, face = "bold", size = rel(1.66)*plot_size_ratio, colour = "black")) +
      theme(axis.line = element_line(colour = "black", linewidth = 1*plot_size_ratio)) +
      theme(axis.ticks.length = unit(0.85*plot_size_ratio, "cm")) +
      theme(axis.ticks.y = element_line(colour = "white", linewidth = 0)) +
      theme(axis.ticks.x = element_line(colour = "#222222")) +
      theme(axis.ticks.length = unit(0.4*plot_size_ratio, "cm")) + 
      theme(axis.title.x = element_text(family = plot_family, size = rel(2)*plot_size_ratio)) +
      guides(fill = guide_legend(override.aes = list(colour = NULL))) +
      guides(fill = guide_legend(nrow = 1, byrow = TRUE))
  if (y.log10) {
    p <- p + scale_y_continuous(trans="log10")
  }
  if (y.log2) {
    p <- p + scale_y_continuous(trans="log2")
  }
  if (add.geom_text) {
    p <- p + geom_text(aes(label=format_values_2(value), y=0), size=0.75, hjust='left')
  }
  if (!plot.axis.text.y) {
    p <- p + theme(axis.text.y=element_blank())
  }
  return(p)
}
```

Function to generate a formatted taxa names from samples sheet.
```{r func_format_sp_name}
# Format for plotting
# See https://yulab-smu.top/treedata-book/faq.html?q=rename#faq-formatting-label
format_sp_name = function(x){
  n = x[["genera"]]
  
  # Add species if present
  if(x[["species"]] != ""){
    n = paste(n, x[["species"]], sep="~")
  }
  
  # Either print with extra info (not italics) or not
  if(x[["extra"]] != ""){
    e = str_replace_all(x[["extra"]], " ", "~")
    e = str_replace_all(e, ",", "")
    lab = glue(paste("italic(",n,")~",e, sep=''))
  } else {
    lab = glue(paste("italic(",n,")", sep=''))
  }
  return(lab)
}
```





# Load datasets into R

Load the cleaned and processed datasets into R for plotting.
```{r load_data}
species.tree <- read.iqtree(
  "../01_Orthofinder/Run2.SpeciesTree_rooted_node_labels.tre")
species.names <- as.phylo(species.tree)$tip.label

samples <- read.csv("../samples.txt", header=TRUE, sep='\t') %>% 
  filter(sample_id %in% species.names)

genome.stats <- read.table(
  "all_genomes-01_stats-results.tsv", sep='\t', 
  header=TRUE, check.names=FALSE, stringsAsFactors=FALSE) %>% 
  filter(SampleID %in% species.names)
transcriptome.stats <- read.table(
  "all_transcriptomes-01_stats-results.tsv", sep='\t', 
  header=TRUE, check.names=FALSE, stringsAsFactors=FALSE) %>% 
  filter(SampleID %in% species.names) %>%
  rename(`Number of genes` = `Number of CDS`)

genomeData.busco.genome.eukaryota <- read.table(
  "all_genomes-02_busco-genome.fa.busco_eukaryota_odb10-results.tsv", sep='\t', 
  header=TRUE, check.names=FALSE) %>% 
  filter(SampleID %in% species.names)
genomeData.busco.genome.metazoa <- read.table(
  "all_genomes-02_busco-genome.fa.busco_metazoa_odb10-results.tsv", sep='\t', 
  header=TRUE, check.names=FALSE) %>% 
  filter(SampleID %in% species.names)
genomeData.busco.protein.eukaryota <- read.table(
  "all_genomes-02_busco-pep.faa.busco_eukaryota_odb10-results.tsv", sep='\t', 
  header=TRUE, check.names=FALSE) %>% 
  filter(SampleID %in% species.names)
genomeData.busco.protein.metazoa <- read.table(
  "all_genomes-02_busco-pep.faa.busco_metazoa_odb10-results.tsv", sep='\t', 
  header=TRUE, check.names=FALSE) %>% 
  filter(SampleID %in% species.names)

transcriptomeData.busco.protein.eukaryota <- read.table(
  "all_transcriptomes-02_busco-pep.faa.busco_eukaryota_odb10-results.tsv", sep='\t', 
  header=TRUE, check.names=FALSE) %>% 
  filter(SampleID %in% species.names)
transcriptomeData.busco.protein.metazoa <- read.table(
  "all_transcriptomes-02_busco-pep.faa.busco_metazoa_odb10-results.tsv", sep='\t', 
  header=TRUE, check.names=FALSE) %>% 
  filter(SampleID %in% species.names)
```

Merge transcriptome and genome datasets together so that all species from the tree are represented (species without this data get zero values).
```{r merge_datasets}
# Genome/Transcriptome stats
seq.stats <- merge(genome.stats, transcriptome.stats, all=TRUE) %>% replace(is.na(.), 0)

# BUSCO results
busco.protein.eukaryota <- rbind(genomeData.busco.protein.eukaryota, transcriptomeData.busco.protein.eukaryota)
busco.protein.metazoa <- rbind(genomeData.busco.protein.metazoa, transcriptomeData.busco.protein.metazoa)

busco.genome.eukaryota <- data.frame(SampleID=species.names)
busco.genome.eukaryota <- merge(x=busco.genome.eukaryota, y=genomeData.busco.genome.eukaryota, all.x=TRUE) %>% replace(is.na(.), 0)

busco.genome.metazoa <- data.frame(SampleID=species.names)
busco.genome.metazoa <- merge(x=busco.genome.metazoa, y=genomeData.busco.genome.metazoa, all.x=TRUE) %>% replace(is.na(.), 0)
```

Extract the sample order from dendrogram (using "working" sample names). Also load pretty sample names and replace the labels in the tree with these new names.





# Generate plot

Global setting for plotting
```{r plot_settings}
plot_size_ratio <- 0.2
plot.axis.text.y <- FALSE
plot.legend <- FALSE
```

Data.frame with clade positions in tree and colors.
```{r clade_colors}
# Dataframe of clades to highlight
clade.colors <- data.frame(
  node  = c(132,        138,         144,         146,       148,        164,            171,            172,          183,            179,                130,        126,        122,          124,                134,       136,          133,             56,          57),
  fill  = c("blue",     "red",       "brown",     "pink",    "yellow",   "purple",       "green",        "orange",     "blue",         "red",             "purple",    "green",    "yellow",     "red",              "blue",    "blue",       "yellow",       "green",        "blue"),
  alpha = c(0.2,        0.4,         0.2,         0.4,       0.2,        0.2,            0.2,            0.2,          0.2,            0.2,                0.5,        0.5,        0.5,          0.5,                0.5,       0.1,          0.2,            0.4,          0.2),
  n     = c("Cnidaria", "Staurozoa", "Scyphozoa", "Cubozoa", "Hydrozoa", "Octocorallia", "Hexacorallia", "Actiniaria", "Scleractinia", "Corallimorpharia", "Placozoa", "Porifera", "Ctenophora", "Choanoflagellata", "Myxozoa", "Medusozoa",  "Endocnidozoa", "Zoantharia", "Scleractinia")
)
```

Format species tree. Color clades and rename taxa into a pretty format (i.e., genera+species italics + extra info not italics).
 - Show tree with full taxonomy to verify that our clades are in the correct position.
```{r species_tree_taxonomy}
samples$name <- apply(samples, 1, format_sp_name )
p <- ggtree(species.tree) %<+% samples + 
  geom_highlight(
    data = clade.colors,
    mapping = aes(
      node=node,
      fill=I(fill),
      alpha=I(alpha),
    ),
    extendto = 1.65,
    to.bottom=TRUE,
  ) +
  geom_cladelab(
    data = clade.colors,
    mapping = aes(
      node=node,
      label=n,
    ),
    align = TRUE,
    offset.text = 0.0,
    hjust = "center",
    fontsize = 2,
    offset = 0.5,
    barsize = 0,
    fontface="bold",
    ) +
  geom_tiplab(aes(label=name),     size=6*plot_size_ratio, linesize=0.2, parse=TRUE, align=TRUE, offset=0.01) +
  geom_tiplab(aes(label=order),    size=6*plot_size_ratio, linesize=0.0, parse=TRUE, align=TRUE, offset=0.18) +
  geom_tiplab(aes(label=subclass), size=6*plot_size_ratio, linesize=0.0, parse=TRUE, align=TRUE, offset=0.25) +
  geom_tiplab(aes(label=class),    size=6*plot_size_ratio, linesize=0.0, parse=TRUE, align=TRUE, offset=0.33) +
  geom_tiplab(aes(label=phylum),   size=6*plot_size_ratio, linesize=0.0, parse=TRUE, align=TRUE, offset=0.40) +
  geom_tippoint(aes(color=data_type), size=6*plot_size_ratio, show.legend=FALSE) +
  hexpand(0.1)

p + geom_text(aes(label=node))
```



Generate data plots that we will combine together in the final plot.
```{r data_plots}
p.genome.stats.scafs <- plot_barchart(data=seq.stats,
                                    column.id="Number of scaffolds", 
                                    plot_title="No. scaffolds", 
                                    xlab="Number of scaffolds", 
                                    bar.fill="#fdb462", 
                                    plot.axis.text.y=plot.axis.text.y,
                                    plot_size_ratio=plot_size_ratio)
p.genome.stats.bases <- plot_barchart(data=seq.stats,
                                    column.id="Total scaffold length (bp)", 
                                    plot_title="Total bases", 
                                    xlab="Bases (Mbp)", 
                                    bar.fill="#fb8072", 
                                    divide.by=1000000,
                                    plot.axis.text.y=plot.axis.text.y,
                                    plot_size_ratio=plot_size_ratio)
p.genome.stats.N50 <- plot_barchart(data=seq.stats,
                                    column.id="N50 of scaffolds (bp)", 
                                    plot_title="Assembly scaffold N50", 
                                    xlab="Scaffold N50 (Kbp)", 
                                    bar.fill="#bebada", 
                                    divide.by=1000,
                                    plot.axis.text.y=plot.axis.text.y,
                                    plot_size_ratio=plot_size_ratio)
p.gene.stats.count <- plot_barchart(data=seq.stats,
                                    column.id="Number of genes", 
                                    plot_title="Protein-coding genes", 
                                    xlab="Number of genes", 
                                    bar.fill="#d9d9d9", 
                                    plot.axis.text.y=plot.axis.text.y,
                                    plot_size_ratio=plot_size_ratio)

p.busco.genome.eukaryota  <- plot_BUSCO_percentages(
                                    data=busco.genome.eukaryota, 
                                    plot_title="BUSCO Eukaryota - Genome", 
                                    plot.axis.text.y=plot.axis.text.y,
                                    plot.legend=plot.legend,
                                    plot_size_ratio=plot_size_ratio)
p.busco.genome.metazoa    <- plot_BUSCO_percentages(
                                    data=busco.genome.metazoa, 
                                    plot_title="BUSCO Metazoa - Genome", 
                                    plot.axis.text.y=plot.axis.text.y,
                                    plot.legend=plot.legend,
                                    plot_size_ratio=plot_size_ratio)
p.busco.protein.eukaryota <- plot_BUSCO_percentages(
                                    data=busco.protein.eukaryota, 
                                    plot_title="BUSCO Eukaryota - Protein", 
                                    plot.axis.text.y=plot.axis.text.y,
                                    plot.legend=plot.legend,
                                    plot_size_ratio=plot_size_ratio)
p.busco.protein.metazoa   <- plot_BUSCO_percentages(
                                    data=busco.protein.metazoa, 
                                    plot_title="BUSCO Metazoa - Protein", 
                                    plot.axis.text.y=plot.axis.text.y,
                                    plot.legend=plot.legend,
                                    plot_size_ratio=plot_size_ratio)
```

Plot final combined panel plot.
```{r plot_QC_metrics_Run1}
samples$name <- apply(samples, 1, format_sp_name )
p <- ggtree(species.tree) %<+% samples + 
  geom_highlight(
    data = clade.colors,
    mapping = aes(
      node=node,
      fill=I(fill),
      alpha=I(alpha),
    ),
    extendto = 1.400, # Change position of colored boxes
    to.bottom=TRUE,
  ) +
  geom_cladelab(
    data = clade.colors,
    mapping = aes(
      node=node,
      label=n,
    ),
    align = TRUE,
    offset.text = 0.0,
    hjust = "center",
    fontsize = 2,
    offset = 0.2, # Push everything to the right
    barsize = 0,
    fontface="bold",
    ) +
  geom_tiplab(aes(label=name), size=6*plot_size_ratio, linesize=0.2, parse=TRUE, align=TRUE, offset=0.01) +
  geom_tippoint(aes(color=data_type), size=6*plot_size_ratio, show.legend=FALSE) +
  hexpand(.1)

p.gene.stats.count %>% 
  insert_left(p, width=6) %>% 
  insert_right(p.genome.stats.scafs) %>% 
  insert_right(p.genome.stats.bases) %>% 
  insert_right(p.genome.stats.N50) %>%
  insert_right(p.busco.genome.eukaryota) %>% 
  insert_right(p.busco.genome.metazoa) %>% 
  insert_right(p.busco.protein.eukaryota) %>%
  insert_right(p.busco.protein.metazoa)
```

Plot legends for BUSCO plot.
```{r plot_QC_legends_Run1}
plot_BUSCO_percentages(
  data=busco.genome.eukaryota, 
  plot_title="BUSCO Eukaryota - Genome", 
  plot.axis.text.y=plot.axis.text.y,
  plot.legend=TRUE,
  plot_size_ratio=plot_size_ratio) %>%
  get_legend() %>%
  as_ggplot()
```





# Session Info

```{r ressionInfo}
sessionInfo()
```
